A fast unified model for parsing and sentence understanding

### 2016-03-15 writing/hard_stack_paper/paper.tex
Hasty pre-NLP Group edits.

\input{batching_fig.tex}

@@

\input{../model1_fig.tex}

@@

In this paper, we introduce the Stack-augmented Parser-Interpreter Neural Network, or SPINN. The SPINN executes tree-structured computations in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the previous TreeRNN approach to compositionality in three ways:
\begin{itemize}
\item At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser.
\item It supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
\item It supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
\end{itemize}

@@

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$. It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. (For a binary-branching tree structure, this requires $T = 2N - 1$ operations.)

The parser uses two auxiliary data structures, a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. We initialize the parser with an empty stack and with a buffer containing all the sentence tokens $\mathbf x$. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. We next proceed through the transition sequence, where each transition $t_i$ selects one of the two following operations. Below $\mid$ denotes an infix \textit{cons} function.
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it onto the top of the stack.
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back onto the stack.
\end{description}

@@

The core SPINN model implements a sequential shift-reduce parser which operates over sentences. It is designed to produce a vector representation of the sentence as output (as opposed to a tree). This is accomplished by the following modifications to the shift-reduce algorithm:
\begin{itemize}
\item The intermediate representations in the parser, including the representations of partial tree structures on the stack, are fixed-length vectors.
\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a neural network.
\end{itemize}

This subsection describes the core functions and representations of the SPINN. We begin with the most basic model, SPINN-NP (for \textit{no parser}), which processes input sentences with gold parses provided by some external parser.

\paragraph{The buffer and word representations}

We draw our word representations from the standard 300-dimensional vector package provided with GloVe \cite{pennington2014glove}. We do not update these vectors during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors that can be used as inputs to a TreeLSTM composition function:

@@

In a SPINN, each row of the stack contains a pair of vectors $(h, c)$, which together represent some node in the parse tree of the sentence. Since the \shift~operation simply copies entries from the buffer onto the stack, this means that the word representations in the buffer also have two parts, $h$ and $c$.

@@

Our composition function is based on the TreeLSTM of \citealt{tai2015improved}. The TreeLSTM generalizes the LSTM neural network layer \cite{hochreiter1997long} to tree- rather than sequence-based inputs, and it shares with that older design the idea of representing intermediate states in a computation using a two-part vector representation containing an $\vec{h}$ vector and a $\vec{c}$ vector, where the latter is meant to serve as a slower-changing long-term memory.  It is formulated as:

@@

The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack. The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$. In addition, $\vec{e}$ is an optional input argument which is either the empty vector $[]$ or a vector from an external source like the tracking LSTM (see below). $\odot$ denotes the elementwise product. Each vector-valued variable listed here is of dimension $D$, except $\vec{e}$, which is of dimensions $D_t$, which can be chosen independently.

@@

This paper presents the transition-based sentence model within the context of \textit{natural language inference} (also known as \textit{recognizing textual entailment}), a sentence pair classification task. To classify a sentence pair, a feature vector is first constructed. This feature vector is based on the final representations of each of the two sentences---the representations that appear at the head of the stack for each sentence after the final transition. In particular, the $\vec{h}$ portions of these representations are used. The final feature vector consists of the concatenation of these two vectors, $\vec{x}_{\text{premise}}$ and $\vec{x}_{\text{hypothesis}}$, their difference, and their elementwise product:

@@

This feature vector is then passed to a series of two (\todo{Update w/ final number}) ReLU neural network layers (following \citealt{snli:emnlp2015}), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.

\subsection{Tracking left context with an auxiliary LSTM}

While it is possible to use the model that we have described thus far as a sentence encoder, at least when an external parser is available to supply \shift~and \reduce~instructions (and we do evaluate such a model), our full model includes an additional component: the tracking LSTM. The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step. It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which is what allows the model to stand alone as a parser, and it additionally supplies a secondary input to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.

Technically, the tracking LSTM takes three vectors as inputs at each step: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the two two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation. Its output hidden state at each step is used as the external input to the TreeLSTM composition function for that step and as the sole input to the transition prediction function.

@@

\todo{[JG] Include figure of post-order traversal sequence made by the tracking LSTM?}
\todo{[SB] Say more about joint learning of phrase representations for the two objectives.}

@@

\subsection{Implementing the transition-based sentence model}

@@

We trained each model for 250k steps in each run, using a minibatch size of 32 for each step. We tracked each model's performance on the development set during training and saving parameters when this performance reached a new peak. We used early stopping, evaluating on the test set using the parameters that performed best on the development set.

@@

A naive implementation of the SPINN model would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a total space requirement of $B \times N \times T \times D$, which is prohibitively large for significant batch sizes $B$ or sentence lengths $N$. Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).

>>>>

\input{batching_fig.tex}
\input{../model1_fig.tex}

@@

@@

In this paper, we introduce the Stack-augmented Parser-Interpreter Neural Network, or SPINN. The SPINN executes tree-structured computations in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the TreeRNN architecture in three ways:
\begin{itemize}
\item At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser.
\item It supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
\item It supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
\end{itemize}

\todo{[SB] Move the introduction of the SPINN earlier in the introduction. Mention that we'll be using SNLI as our sole evaluation task.}

@@

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$. It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions.

The parser uses two auxiliary data structures, a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. We initialize the parser with an empty stack and with a buffer containing all the sentence tokens $\mathbf x$. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. We next proceed through the transition sequence, where each transition $t_i$ selects one of the two following operations. Below $\mid$ denotes an infix \textit{cons} function. \todo{[JG] Clarify prev. sentence.}
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it onto the top of the stack.
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back onto the stack.
\end{description}

@@

The core SPINN model implements a sequential shift-reduce parser which operates over sentences. It is designed to produce a vector representation of the sentence as output,as opposed to a tree as in standard shift-reduce \textit{parsing}. This relies on two key modifications to the shift-reduce algorithm:
\begin{itemize}
\item The parser's intermediate representations, including the representations of partial tree structures on the stack, are fixed-length vectors.
\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a neural network.
\end{itemize}

This subsection describes the core functions and representations of the SPINN. We begin with the most basic model, SPINN-NP (for \textit{no parser}), which processes input sentences using parses provided by some external parser.

\paragraph{Word representations}

The buffer is populated with the words in each sentence before the first step of computation. We use word representations from the standard 300-dimensional vector package provided with GloVe \cite{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors that are stored side-by-side in the buffer and used as inputs to a TreeLSTM composition function:

@@

In the SPINN, each row of the stack and each row of the buffer contains a pair of vectors $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence. \todo{[SB] Clarify that this is needed because of the TreeLSTM fn.}

@@

Our composition function is based on the TreeLSTM of \citealt{tai2015improved}. The TreeLSTM generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with that older design the idea of representing intermediate states in a computation using a vector pair $\langle \vec{h}, \vec{c}\rangle$, of which the latter vector is meant to serve as a slower-changing long-term memory.  It is formulated as:

@@

The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack. The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$. In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see below). $\odot$ denotes the elementwise product. Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.

@@

This paper evaluates the SPINN within the context of \textit{natural language inference} (also known as \textit{recognizing textual entailment}), a sentence pair classification task. To classify a sentence pair, a feature vector is first constructed. This feature vector is based on the final representations of each of the two sentences---the representations that appear at the head of the stack for each sentence after the final transition. In particular, the $\vec{h}$ portions of these representations are used. The final feature vector consists of the concatenation of these two vectors, $\vec{x}_{\text{premise}}$ and $\vec{x}_{\text{hypothesis}}$, their difference, and their elementwise product:

@@

This feature vector is then passed to a series of two (\todo{[SB] Update number}) ReLU neural network layers (following \citealt{snli:emnlp2015}), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels. The model is trained in part using a cross-entropy objective over this distribution.

\subsection{The tracking LSTM}

While it is possible to use the model that we have described thus far as a sentence encoder, at least when an external parser is available to supply \shift~and \reduce~instructions (and we do evaluate such a model), our full model includes an additional component: the tracking LSTM. (\todo{[SB] Update framing of prev sentence.}) The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step. It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which is what allows the model to stand alone as a parser, and it additionally supplies a secondary input to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.

Technically, the tracking LSTM takes three vectors as inputs at each step: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation. Its output hidden state at each step is used as the external input to the TreeLSTM composition function for that step and as the sole input to the transition prediction function.

@@

\todo{[JG] Include figure of post-order traversal sequence made by the tracking LSTM?}

@@

\subsection{Implementing the SPINN}

\todo{[SB] Say more about joint learning of phrase representations for the two objectives.}

@@

We trained each model for 250k steps in each run, using a batch size of 32 for each step. We tracked each model's performance on the development set during training and saving parameters when this performance reached a new peak. We used early stopping, evaluating on the test set using the parameters that performed best on the development set.

@@

A naive implementation of the SPINN model would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a total space requirement of $M \times N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$. Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).
